---
title: "hw6"
author: "mhanyu"
format:
  html:
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
---
[My Github](https://github.com/Mahanyuu/STAT-506/)

## PS1

### 1.a

```{r 1}
library(DBI)
lahman <- dbConnect(RSQLite::SQLite(), "../data/lahman_1871-2022.sqlite")

gg <- function(query){
  dbGetQuery(lahman, query)
}

fielding <- gg("
   SELECT * FROM Fielding
   ")
dbDisconnect(lahman)
```


```{r 1.a.0}
library(dplyr)
fielding <- fielding[fielding$InnOuts!=0, ]
fielding$RF <- with(data = fielding, 3*(PO+A)/InnOuts)

## mean
RF_mean <- fielding %>%
  group_by(teamID)%>%
  summarize(mean_RF = mean(RF, na.rm=TRUE)) %>%
  filter(is.finite(mean_RF))%>%
  arrange(-mean_RF)
RF_mean
```

#### 1.a.1 Without any parallel processing
```{r 1.a.1}
## sd estimation 
### Without any parallel processing
strata <- unique(fielding$teamID)
nsim <- 2

#' Title return the mean of bootstrap resample per stratum
#'
#' @param strata input the unique strata to the function
#' @param data input the data frame
#'
#' @return the mean of bootstrap resample per stratum


boot_strata <- function(strata, data){
  n <- length(strata)
  bootstrap_sample <- data.frame(teamID = character(n), mean_RF = numeric(n))
  
  for (i in seq_along(strata)){
    stratum <- strata[[i]]
    stratum_data <- data[data$teamID==stratum, ]
    resample <- stratum_data[sample(nrow(stratum_data), replace=TRUE), ]
    bootstrap_sample[i, ] <- c(stratum, mean(resample$RF, na.rm =TRUE))
  }
  
  return(bootstrap_sample)
}

#lapply
set.seed(789)
resc_1 <- lapply(seq_len(nsim), function(x)boot_strata(strata, fielding))
resc_1_combined <- do.call(rbind, resc_1)
resc_1_combined$mean_RF <- as.numeric(resc_1_combined$mean_RF)
resc_1_all <- resc_1_combined %>%
  group_by(teamID) %>%
  summarize(mean_1 = mean(mean_RF, na.rm=TRUE), sd_1=sd(mean_RF, na.rm=TRUE))
resc_1_all
```

#### 1.a.2 Using parallel processing with the parallel package.
```{r 1.a.2}
### Using parallel processing with the parallel package.
library(parallel)
cl <- makeCluster(detectCores()-1)
clusterExport(cl, varlist=c("strata", "fielding", "boot_strata"))
set.seed(678)
resc_2 <- parLapply(cl, seq_len(nsim), function(x)boot_strata(strata, fielding))
stopCluster(cl)
resc_2_vector <- do.call(rbind, resc_2)
resc_2_clean <- resc_2_vector[is.finite(as.numeric(resc_2_vector$mean_RF)), ]
resc_2_all <- resc_2_clean %>%
  group_by(teamID) %>%
  summarize(mean_2 = mean(mean_RF), sd_2 = sd(mean_RF))
```

#### 1.a.3 Using futures with the future package.
```{r 1.a.3}
### Using futures with the future package.
library(future)
library(furrr)

# Set up a plan for parallel computation
plan(multisession)  # Suitable for Windows
set.seed(567)
# Use future-powered lapply
resc_3 <- suppressWarnings(future_map(seq_len(nsim), function(x) boot_strata(strata, fielding)))
resc_3_vector <- do.call(rbind, resc_3)
resc_3_clean <- resc_3_vector[is.finite(as.numeric(resc_3_vector$mean_RF)), ]
resc_3_all <- resc_3_clean %>%
  group_by(teamID) %>%
  summarize(mean_3 = mean(mean_RF), sd_3 = sd(mean_RF))
```

### 1.b
```{r 1.b}
team_top_10 <-RF_mean[1:10, ]$teamID
resc_1_2 <- merge(resc_1_all, resc_2_all, by = "teamID", all.x = TRUE)
resc_all <- merge(resc_1_2, resc_3_all, by = "teamID", all.x=TRUE)
resc_all_top_10 <- resc_all[resc_all$teamID %in% team_top_10, ]
resc_all_top_10_order <- resc_all_top_10 %>%
  arrange(match(teamID, team_top_10))
resc_all_top_10_order
```

### 1.c
```{r 1.c}
system.time({
  set.seed(124)
  resc_1 <- lapply(seq_len(nsim), function(x)boot_strata(strata, fielding))
})

system.time({
  cl <- makeCluster(detectCores()-1)
  clusterExport(cl, varlist=c("strata", "fielding", "boot_strata"))
  set.seed(234)
  resc_2 <- parLapply(cl, seq_len(nsim), function(x)boot_strata(strata, fielding))
  stopCluster(cl)
})

system.time({
  plan(multisession)  # Suitable for Windows
  set.seed(123)
  # Use future-powered lapply
  resc_3 <- suppressWarnings(future_map(seq_len(nsim), function(x) boot_strata(strata, fielding)))
})
```
It is very obvious that execution time with lapply is the longest. With the future and parallel packages help, the time is shortened effectively. For the elapsed total time, the program without any parallel processing runs 664.86s. Using parallel processing, the program runs 167.25s. Using futures with future package, the program runs 164.86s. The program with parallel processing runs fast, and the second is future processing. They are both further faster than processing without any parallel processing.
